### 2010-06-29 trunk/paper/IIWAS2010/content/related_work.tex
edit section 2.2

% This section needs to be reworkd. It should be about 1 paragraph. It should define
% what these (ontology, dbpedia, wordnet) mean to morpheus
% What rigors boone invest to my pyramids? <<-- randomly generated

%An ontology formally models real world \textit{concepts} and theirrelationships. A concept or class in an ontology gives us an abstract and simplified view of the world\cite{Gruber1993} in a machine-readable and language-independent format. The relationships and attributes of an ontological concept are defined using \textit{properties} and \textit{property restrictions}. An ontology definition contains all of these: classes, properties, and restrictions. A class in an ontology can have multiple \textit{super classes} and \textit{sub-classes}. Finally, a \textit{knowledge base} represents an ontology together with instances of its classes.

@@

The DBpedia\footnote{http://dbpedia.org} project is based on extracting  semantic information from the Wikipedia and making it available on the Web.  Wikipedia semantics includes info-box templates, categorization information,  images, Geo-coordinates, links to external Web pages, disambiguation pages,  and redirects between pages in Wiki markup form \cite{Bizer2009}.  DBpedia represents the Wikipedia categories using skos:concepts and category relations using skos:broader \cite{Bizer2009}. In fact,  DBpedia does not define any new relations between the Wikipedia categories.  The Wikipedia categories are extracted directly from Wikipedia pages, and there  is no quality check on the resultant ontology. Similarly, additional ontological relations in DBpedia are generated from the info-box attributes and their values.

YAGO is a semi-automatically constructed ontology from the Wikipedia pages, info-boxes, categories and WordNet\footnote{http://wordnet.princeton.edu} synsets heterarchy\cite {Suchanek2009phd}. YAGO uses the Wikipedia page titles as its ontology individuals and categories as its ontology classes. The ontology's heterarchy is built using the \textit{hypernym} and \textit{hyponym} relations of the WordNet synsets. YAGO uses only the nouns from WordNet and ignores the WordNet verbs and adjectives. The connection between a WordNet synset and a Wikipedia category is achieved by parsing the category names and matching the parsed category components with the WordNet synsets\cite{Suchanek2009phd}. Those Wikipedia categories having no WordNet match are ignored in the YAGO ontology.

However, YAGO type extraction assumes that each Wikipedia page has at least one tagged category, and the assigned Wikipedia categories are relevant to that Wikipedia page. The Wikipedia page category assignments are subjectively assigned by human editors and some Wiki pages are unassigned i.e., they have no category. In addition, we cannot fully rely on the relevance of these assigned Wikipedia page categories. In addition, YAGO concentrates mainly on the fields of people, cities, events, and movies\cite{Suchanek2009phd}.

>>>>

% This section needs to be reworked. It should be about 1 paragraph. It should define
% what these (ontology, dbpedia, wordnet) mean to morpheus
% What rigors boone invest to my pyramids? <<-- randomly generated

%An ontology formally models real world \textit{concepts} and their relationships. A concept or class in an ontology gives us an abstract and simplified view of the world\cite{Gruber1993} in a machine-readable and language-independent format. The relationships and attributes of an ontological concept are defined using \textit{properties} and \textit{property restrictions}. An ontology definition contains all of these: classes, properties, and restrictions. A class in an ontology can have multiple \textit{super classes} and \textit{sub-classes}. Finally, a \textit{knowledge base} represents an ontology together with instances of its classes.

@@

The DBpedia\footnote{http://dbpedia.org} project is based on a community of contributors extracting semantic information from the Wikipedia and making it available on the Web.  Wikipedia semantics includes info-box templates, categorization information,  images, Geo-coordinates, links to external Web pages, disambiguation pages, and redirects between pages in Wiki markup form \cite{Bizer2009}.  DBpedia does not define any new relations between the Wikipedia categories.  Morpheus the DBpedia categories and its semantics to construct its ontology.
%  DBpedia represents the Wikipedia categories using skos:concepts and category relations using skos:broader \cite{Bizer2009}. In fact,  DBpedia does not define any new relations between the Wikipedia categories.  The Wikipedia categories are extracted directly from Wikipedia pages, and there  is no quality check on the resultant ontology. Similarly, additional ontological relations in DBpedia are generated from the info-box attributes and their values.

YAGO is a semi-automatically constructed ontology from the Wikipedia pages, info-boxes, categories and WordNet\footnote{http://wordnet.princeton.edu} synsets heterarchy \cite{Suchanek2009phd}. YAGO uses the Wikipedia page titles as its ontology individuals and categories as its ontology classes. YAGO uses only the nouns from WordNet and ignores the WordNet verbs and adjectives.  YAGO discovers connections between a WordNet synsets and a Wikipedia categories parsing the category names and matching the parsed category components with the WordNet synsets \cite{Suchanek2009phd}. Those Wikipedia categories having no WordNet match are ignored in the YAGO ontology.  The ontology's heterarchy is built using the \textit{hypernym} and \textit{hyponym} relations of the WordNet synsets.  Morpheus uses many of the same ideas made when using YAGO to develop its own ontology as explained in Section \ref{sec:ontology_corpora}.

%However, YAGO type extraction assumes that each Wikipedia page has at least one tagged category, and the assigned Wikipedia categories are relevant to that Wikipedia page. The Wikipedia page category assignments are subjectively assigned by human editors and some Wiki pages are unassigned i.e., they have no category. In addition, we cannot fully rely on the relevance of these assigned Wikipedia page categories. In addition, YAGO concentrates mainly on the fields of people, cities, events, and movies\cite{Suchanek2009phd}.

### 2010-06-29 trunk/paper/IIWAS2010/content/system_arch.tex
edit section 2.2

This section presents the core elements of the Morpheus system and its 
role in the \textit{automated question answering} process.

\subsection{Query Processing}
\label{sec:query_processing}

The \emph{Query Resolution Recorder} or QRR is an instrumented web browser
used to track the search pathways to an answer for a \textit{guide's} query. It
also helps guide to identify categories (optional) for the query terms. 
Morpheus has two types of categories: \emph{intput categories} and \emph{output categories}.
Input categories annotates the query terms into a class or category, which are 
useful in finding similar queries. On the other hand, output categories give hints
to the expected answer of a user query. Morpheus captures all
this query information in the representation of an SSQ. For example, suppose, a guide 
tries to answer the query ``A 1997 Toyota Camry V6 needs what size tires?'', using
Morpheus. With the help of the QRR, guide may choose relevant query terms and assign
\emph{input categories} to them: Toyota : manufacturer, Camry V6 : 
model, tire : automotive parts, and size as the \emph{output category}. 
In addition, guide can log the search pathways to an answer (e.g. P215/60R16)
to this query and select the query \emph{realm} as automotive.

The \emph{Query Resolution Method} or QRM is a data structure that models the
question answering process. A QRM is usually constructed by a \emph{guide} with
the help of the QRR. Each QRM contains an \emph{ontological realm}, an SSQ, and
information to support the query answering process. QRMs are associated with an
ontology that has a particular realm i.e. an ontological realm. In addition, 
the QRM contains information required to visit each page needed to 
answer the query as well as the order of visitation. For each
dyanamic page, it stores the list of inputs to the URL such as form inputs and
the possible referrence outputs (highlighted portions of the web page).

@@

Rather than working blindly, forming all possible ngrams from the query 
and running probabilities, it seemed to be a better approach to identify 
important noun phrases within the query, and work with those. Using our 
query parser, we proceed to tokenize, parse, and chunk our candidate query, 
extracting verb groups, answer classes, and descriptive info from each query. 
In the example query "What is the tire size of a 1997 Toyota Camry?," 
the verb group is the verb "is," the answer class is "tire size," and descriptive 
info is "of a 1997 Toyota Camry." The next step is to create ngrams from these chunks. 
Using these ngrams, we will query our database for frequencies associated with 
each ngram: overall frequency and categorical frequency, alongside category 
term count and overall term count, to calculate categorical probabilities 
and establish a realm for the candidate query, choosing realm with highest probability.

\subsection{Morpheus Ontology and Corpora} 
\label{sec:ontology_corpora}

Morpheus requires a realm-based ontology that contains categories of a
particular realm of interest. For example, the realm-ontology \textit{Automotive}
contains categories that are relevant to the automotive realm. In
addition, every leaf node in the category ontology is associated with a corpus of words
that belongs to that category.

For ontology construction Morpheus uses the Wikipedia pages, categories,
and the WordNet synset heterarchy. This approch is motivated from
YAGO\cite{Suchanek2009phd}. For a given realm, we first find a mapping category 
from the DBpedia categories\cite{Bizer2009}. Then we grab all the neighboring 
categories using a \textit{Markov Blanket} \cite{PRIS}. In fact, we grab 
the category's parents, its children, and its 
children's other parents using the DBpedia ontology properties \textit{broader}
and \textit{narrower}. Once the blanket is ready, it grabs all parent categories
of the categories in the blanket till it reaches root. Once we have the
categories belongs to a particular domain, we associate them with WordNet synsets
using an algorithm similar to YAGO\cite{Suchanek2009phd}. We determine head noun in the
category name and search for it in the WordNet synsets.
Those Wikipedia categories having no WordNet match are ignored in the our realm
ontology. In addition, we uses the WordNet synset heterarchy relationships for building the ontologcial structure.

To build a corpus for each of the leaf nodes in the ontology, 
we extract \emph{n-grams} or \emph{terms} from the Wikipedia pages associated with the 
the DBpedia categories\cite{Bizer2009}. From this category-term corpora, 
we can find out the likelihood that a term belongs to a particular category, 
which is useful in categorizing the terms in a \emph{regular user's} query [section \ref{sec:qrm_ranking}].    
This likelihood is determined by the probability of a category given a term 
using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can 
easily obtain the term-category and term-corpus probabilities as 
relative frequencies. The higher the probability, the more likely a 
term is referring to a category.

Alternately, we are evaluating Latent Dirichlet Allocation (LDA), a probabilistic graphic model, for
representing latent topics of the documents in a text corpus\cite{Blei2003latentdirichlet}.
Due to its fully generative semantics, even at 
the level of documents, LDA could address the drawbacks 
(such as dimensionality and failure in finding the discriminative set of 
words for a document) of frequency based approaches such as TF-IDF, LSI, and pLSI. In
this Bayesian model, LDA represents a document in the corpus by
distributions over topics, and a topic itself is a distribution over all 
terms in the corpus.  For example, for the Wikipedia pages, the latent
topics reflect the thematic structure of the pages. Thus, LDA
can find out relevant topic proportions in a document using posterior inference
\cite{Blei2003latentdirichlet}. Additionally, given a text document, using LDA
we could tag related documents by matching the similarity over the estimated
topic proportions, which is useful in the ontology and corpora building.

\subsection{QRM Ranking Mechanism} 
\label{sec:qrm_ranking}

@@

We consider the category match as a type match and we use 
\emph{category divergence} to calculate the relevance between 
the candiate SSQ and a qualified SSQ. Each qualified SSQ will 
have input terms, output terms, and associated classes, and one realm (from QRM). 
For the candidate SSQ, the relevant categories for terms 
are determined from the NLP engine and corpora. The calculation of a realm 
for a candidate query is done using the ngrams found within said query 
and the probabilities (if any) found using frequencies within the database.
We match QRMs that belong to the same \emph{realm} of the candidate SSQ. 
The relevance of a qualified SSQ to the candidate SSQ is 
determined by aggregating the divergence measure of input-term-categories 
associated with them. In addition, we order QRMs in the data store 
by descreasing relevance. The order provides a ranking for the results 
to the user. The following subsection describes \emph{category divergence} 
in detail.

@@

Suppose we want to find the category divergence between \textit{Sedans}
and \textit{Trucks} (Figure \ref{fig:automotive_ontology}). 
\textit{Sedans} is the subcategory of \textit{Cars}, and \textit{Trucks}
is the subcategory of \textit{Automobiles}. Therefore, the least common ancestor ($C$)
for these two categories is \textit{Automobiles}. In addition, the tree height \textit{h = 4}.
So, the normalized divergence $cd$\textit{(Sedans, Trucks)} is \textit{7/12}
that is calculated from \textit{d(Sedans, Root) = 4}, \textit{d(Sedans, Automobiles) = 2}, and \textit{d(Trucks, Automobiles) = 1}.

\subsection{Query Resolution Execution}

Usually, answering a question using the 
deep web requires one to navigate through a sequence of one or 
more web pages. Many of the accesses involve clicks through 
web forms or links to resulting pages. QRM contains the necessary  
information required to re-run this procedure. Once we have 
relevant QRMs through QRM ranking for a given user query, we can  
get answers by re-running the pathways with help of 
the Query Resolution Engine or QRE of Morpheus.

>>>>

This section presents ontology and copora, query processing, ranking queries, and query executing.

\subsection{Using Ontology and Corpora} 
\label{sec:ontology_corpora}

Morpheus uses an ontology that contains categories of a particular realm of interest. Every leaf node in the ontology is associated with a corpus of words belonging to a category.  For example, we have constructed an automotive ontology containing categories relevant to the automotive realm. This ontology provides a structure for reference in the following sections.

Morpheus uses the DBpedia categories, Wikipedia pages and the WordNet synset heterarchy to construct ontologies. This approch is motivated by YAGO \cite{Suchanek2009phd}. First a realm is mapped to a DBpedia category \cite{Bizer2009}. 
Then a \emph{Markov Blanket} \cite{PRIS} is created covering all neighboring categories and using the DBpedia ontology properties \emph{broader} and \emph{narrower}. From here, the system recursively includes all categories up to the root.

Once we have the categories belonging to a particular realm, we associate them with WordNet synsets using an approach similar to YAGO\cite{Suchanek2009phd}. We determine a \emph{head noun} in the category name and search for it in the WordNet synsets. DBpedia categories having no WordNet synset match are excluded from our ontology.

To build a corpus for each of the leaf nodes in the ontology, 
we extract \emph{terms} from the Wikipedia pages associated with the 
the DBpedia categories. From this term corpus, 
we can find the likelihood of a term belonging to a particular category. This assists in categorizing terms in a \emph{path-follower} query.
The likelihood is determined by the probability of a category given a term 
using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can 
easily obtain the term-category and term-corpus probabilities as 
relative frequencies.

In addition, we evaluate Latent Dirichlet Allocation (LDA) to identify latent topics of the documents in a corpus \cite{Blei2003latentdirichlet}.  LDA is Bayesian model represents a document in the corpus by distributions over topics, and a topic itself is a distribution over all  terms in the corpus.  For example, for the Wikipedia pages, the latent topics reflect the thematic structure of the pages. Thus, LDA can discover relevant topic proportions in a document using posterior inference \cite{Blei2003latentdirichlet}. Additionally, given a text document, using LDA we could tag related documents by matching the similarity over the estimated topic proportions, which is useful in the ontology and corpora building.  Due to its fully generative semantics, even at the level of documents, LDA could address the drawbacks (such as dimensionality and failure in finding the discriminative set of words for a document) of frequency based approaches such as TF-IDF, LSI, and pLSI.

\subsection{ Recording}
\label{sec:query_processing}

The \emph{Query Resolution Recorder} (QRR) is an instrumented web browser
used to track the search pathways to an answer for a \textit{path-finders} query. It
also helps guide to identify categories (optional) for the query terms. 
Morpheus has two types of categories: \emph{input categories} and \emph{output categories}.
Input categories annotates the query terms into a class or category, which are 
useful in finding similar queries. On the other hand, output categories give hints
to the expected answer of a user query. Morpheus captures all
this query information in the representation of an SSQ. For example, suppose, a guide 
tries to answer the query ``A 1997 Toyota Camry V6 needs what size tires?'', using
Morpheus. With the help of the QRR, guide may choose relevant query terms and assign
\emph{input categories} to them: Toyota : manufacturer, Camry V6 : 
model, tire : automotive parts, and size as the \emph{output category}. 
In addition, guide can log the search pathways to an answer (e.g. P215/60R16)
to this query and select the query \emph{realm} as automotive.

The \emph{Query Resolution Method} (QRM) is a data structure that models the
question answering process. A QRM is usually constructed by a \emph{guide} with
the help of the QRR. Each QRM contains an \emph{ontological realm}, an SSQ, and
information to support the query answering process. QRMs are associated with an
ontology that has a particular realm i.e. an ontological realm. In addition, 
the QRM contains information required to visit each page needed to 
answer the query as well as the order of visitation. For each
dyanamic page, it stores the list of inputs to the URL such as form inputs and
the possible referrence outputs (highlighted portions of the web page).

@@

Rather than working blindly, forming all possible n-grams from the query 
and running probabilities, it seemed to be a better approach to identify 
important noun phrases within the query, and work with those. Using our 
query parser, we proceed to tokenize, parse, and chunk our candidate query, 
extracting verb groups, answer classes, and descriptive info from each query. 
In the example query "What is the tire size of a 1997 Toyota Camry?," 
the verb group is the verb "is," the answer class is "tire size," and descriptive 
info is "of a 1997 Toyota Camry." The next step is to create n-grams from these chunks. 
Using these n-grams, we will query our database for frequencies associated with 
each n-gram: overall frequency and categorical frequency, alongside category 
term count and overall term count, to calculate categorical probabilities 
and establish a realm for the candidate query, choosing realm with highest probability.

\subsection{Ranking} 
\label{sec:qrm_ranking}

@@

We consider the category match as a type match and we use 
\emph{category divergence} to calculate the relevance between 
the candiate SSQ and a qualified SSQ. Each qualified SSQ will 
have input terms, output terms, and associated classes, and one realm (from QRM). 
For the candidate SSQ, the relevant categories for terms 
are determined from the NLP engine and corpora. The calculation of a realm 
for a candidate query is done using the n-grams found within said query 
and the probabilities (if any) found using frequencies within the database.
We match QRMs that belong to the same \emph{realm} of the candidate SSQ. 
The relevance of a qualified SSQ to the candidate SSQ is 
determined by aggregating the divergence measure of input-term-categories 
associated with them. In addition, we order QRMs in the data store 
by descreasing relevance. The order provides a ranking for the results 
to the user. The following subsection describes \emph{category divergence} 
in detail.

@@

Suppose we want to find the category divergence between \textit{Sedans}
and \textit{Trucks} (Figure \ref{fig:automotive_ontology}). 
\textit{Sedans} is the subcategory of \textit{Cars}, and \textit{Trucks}
is the subcategory of \textit{Automobiles}. Therefore, the least common ancestor ($C$)
for these twoThis section presents ontology and copora, Query processing, ranking queries, and the query resolution mechanism. categories is \textit{Automobiles}. In addition, the tree height \textit{h = 4}.
So, the normalized divergence $cd$\textit{(Sedans, Trucks)} is \textit{7/12}
that is calculated from \textit{d(Sedans, Root) = 4}, \textit{d(Sedans, Automobiles) = 2}, and \textit{d(Trucks, Automobiles) = 1}.

\subsection{Executing}

QRM contains the necessary  
information required to re-run this procedure.Once we have 
relevant QRMs through QRM ranking for a given user query, we can  
get answers by re-running the pathways with help of 
the \emph{Query Resolution Engine} (QRE) of Morpheus.
